Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-08 ind_files/03_field_eqns.tex
Remove SSWM line because of confusing terminology in existing literature

The first two assumptions are sometimes called the `weak mutation' limit, and the last is sometimes called the `strong selection' limit, both for obvious reasons.

@@

Our `weak mutation' assumptions imply that the population will be concentrated in an infinitesimal neighborhood around the mean value $y(t)$ (\emph{i.e} that the distribution of traits in the population is sharply peaked). We can thus Taylor expand $w\left(x|N_K\delta_{y(t)}\right)$ about $y(t)$ as:

@@

Note that by definition, $B(y(t))$ is the variance of the trait in the population at time $t$. The term $w\left(z|N_K\delta_{y(t)}\right)$ is the expected growth rate of an individual with trait value $z$ in a population of size $N_K$ in which (almost) every individual has trait value $y$. This quantity is referred to as the `invasion fitness' of a `mutant' trait $z$ in a population of `resident' $y$ individuals. Equation \eqref{AD_canonical_derived} is the canonical form of a broad class of systems called `gradient equations' or `gradient dynamics' in quantitative genetics \citep{lande_natural_1976, abrams_relationship_1993, lehtonen_price_2018, lion_theoretical_2018}, and captures the approximate evolutionary dynamics of quantitative traits under certain mutation limits.  It is also deeply related \citep{lehtonen_price_2018, lion_theoretical_2018} to the canonical equation of adaptive dynamics \citep{dieckmann_dynamical_1996, doebeli_adaptive_2011}. The major difference is that in the `proper' canonical equation of adaptive dynamics (as formulated in \cite{dieckmann_dynamical_1996}), the function $B(y)$ explicitly relies on mutations as a continual source of variation, whereas in gradient dynamics and our equation \eqref{AD_canonical_derived}, $B(y)$ captures the standing genetic variation in the population but does not specify the source of this variation. Note that strictly speaking, if $\psi(x,t) = \delta_{y(t)}$ exactly, then $B(y) \equiv 0$. This just reflects our assumption that mutations are vanishingly rare and mutants are sampled from infinitesimally close to the resident value. More detailed mathematical arguments are required to ensure that this convergence `makes sense' and that $B(y)$ does not actually equal 0. This has been proved rigorously using much more sophisticated mathematical tools grounded in martingale theory \citep{champagnat_unifying_2006}. A heuristic derivation of the canonical equation of adaptive dynamics is provided in the classic article by \cite{dieckmann_dynamical_1996}.

>>>>

@@

Our assumptions on mutation rate and mutational effects imply that the population will be concentrated in an infinitesimal neighborhood around the mean value $y(t)$ (\emph{i.e} that the distribution of traits in the population is sharply peaked). We can thus Taylor expand $w\left(x|N_K\delta_{y(t)}\right)$ about $y(t)$ as:

@@

Note that by definition, $B(y(t))$ is the statistical variance of the trait in the population at time $t$. The term $w\left(z|N_K\delta_{y(t)}\right)$ is the expected growth rate of an individual with trait value $z$ in a population of size $N_K$ in which (almost) every individual has trait value $y$. This quantity is referred to as the `invasion fitness' of a `mutant' trait $z$ in a population of `resident' $y$ individuals. Equation \eqref{AD_canonical_derived} is the canonical form of a broad class of systems called `gradient equations' or `gradient dynamics' in quantitative genetics \citep{lande_natural_1976, abrams_relationship_1993, lehtonen_price_2018, lion_theoretical_2018}, and captures the approximate evolutionary dynamics of quantitative traits under certain mutation limits.  It is also deeply related \citep{lehtonen_price_2018, lion_theoretical_2018} to the canonical equation of adaptive dynamics \citep{dieckmann_dynamical_1996, doebeli_adaptive_2011}. The major difference is that in the `proper' canonical equation of adaptive dynamics (as formulated in \cite{dieckmann_dynamical_1996}), the function $B(y)$ explicitly relies on mutations as a continual source of variation, whereas in gradient dynamics and our equation \eqref{AD_canonical_derived}, $B(y)$ captures the standing genetic variation in the population but does not specify the source of this variation. Note that strictly speaking, if $\psi(x,t) = \delta_{y(t)}$ exactly, then $B(y) \equiv 0$. This just reflects our assumption that mutations are vanishingly rare and mutants are sampled from infinitesimally close to the resident value. More detailed mathematical arguments are required to ensure that this convergence `makes sense' and that $B(y)$ does not actually equal 0. This has been proved rigorously using much more sophisticated mathematical tools grounded in martingale theory \citep{champagnat_unifying_2006}. A heuristic derivation of the canonical equation of adaptive dynamics is provided in the classic article by \cite{dieckmann_dynamical_1996}.
